\acknowledgments

They say it takes a village to raise a child. In much the same way, it takes a village to get a PhD. I am extremely grateful for my ``village" of supporters, including my mentors, collaborators, friends, and family.

I am very thankful to my advisor, Yoav Gilad. He has been extremely influential on me as a scientist, teaching me how to creatively solve challenging problems in human genetics with rigor and focus. His support along the path to graduation has been critical. 

I am thankful to my committee members, Marcelo Nobrega, Matthew Stephens, and Oni Basu. I am also grateful for the opportunity to work with top eating disorder specialists, including Jennifer Wildes, Jessica Baker, and Cindy Bulik. This work would not have been possible without the Human Genetics
community in Cummings Life Science Center. Additionally, I have enjoyed working with with Chris Porras, Dan Rice, Eric Friedlander, and Roy Morgan on the Computational STEM Lab (CSL) outreach program. 

Throughout my PhD, I have been financially supported by the Genetics and Regulation Training Grant (T32GM007197) from the National Institutes
of Health, the National Science Foundation (Graduate Research Fellowship and Innovation Corps), and the University of North Carolina Center of Excellence for Eating Disorders. John Novembre also supported the CSL. Sue Levison started as my ``grants person" and evolved into a friend. Sue made sure that I didn't run away during my first week of graduate school, and has seen me all the way through graduation.   

After about 2 weeks of my rotation in the Gilad lab, I knew that this was where I wanted to do my thesis research. After 5 years, I am still confident that I made the right decision. This feeling is in large part due to my fantastic labmates, past and present. In particular, thank you to Nick Banovich, Seb Pott, Genevieve Housman, Kenneth Barr, Benjamin Fair, Reem Elorbany, Katie Rhodes, Sidney Wang, Po-Young Tung, John Blischak, Briana Mittleman, Ittai Eres, Natalia Gonzales, Jonathan Burnett, Emily Davenport, and Samantha Thomas. Michelle Ward has influenced my scientific thinking tremendously. I am also thankful to the technicians in the lab, particularly Amy Mitrano,
Marsha Myrthil, and Claudia Chavarria, who also helped with my thesis work. I'd also like to thank my close collaborators, particularly Julien Roux and Bryan Pavlovic, without whom Chapters 2 and 3 in this thesis would not have been possible. Joyce Hsiao and Abhishek Sarkar also provided key statistical advice for my work. 

Thank you to my friends Unjin Lee, Charlie Lang, Marissa Pelot, Hillary Childs, Sierra Smart, Sam Lohnes, Tatiana Birgisson, Robin Schwartzman, Janet Scognamiglio, and Aarti Venkat. Marcus Soliai is the best coffee buddy anyone could ask for. It is an honor to call Manny Vazquez my friend and get to be on his ``SACNAS'' team. Finally, it would be impossible to list all the ways that Bryce van de Geijn supported me during my Phd. 


My family has been incredibly supportive throughout my life. In particular, my grandparents, Nannie and Papa, encouraged me via weekly phone calls. Emily and Kate Blake are both amazing people, and I am lucky to call them my sisters.  Thank you to my Mom and Dad, who have loved and encouraged me during my entire life. 